---
format: typst
---

# Lab Worksheet - Exploratory Data Analysis in airquality Dataset

```{r}
#| warning: false
#| message: false

# load libraries:
library(tidyverse) # Recall that tidyverse contains dplyr, ggplot2, and many other useful packages.

# Load `airquality` dataset:
data("airquality")
airquality |> 
    head()
```

Q1: Render this quarto file to a pdf.

\

Q2: Create a summary table of the number and proportion of NA values in each column.

```{r}
# your code here

```

\

Q3: Convert the `Month` column to a factor with it's levels labeled as month names (e.g. "May", "June", etc.).

```{r}
# your code here

```

\

Q4: Create a new column `Temp_C` that converts the temperature from Fahrenheit to Celsius.

```{r}
# your code here

```

\

Q5: Rename the `Wind` column to `Wind_mph` and the Temp column to `Temp_F`.

```{r}
# your code here

```

\

Q6: Create a scatter plot of `Ozone` vs `Temp_C`, colored by `Month`. Add appropriate axis labels and a title.

```{r}
# your code here

```

\

Q7: Create a violin plot of `Ozone` levels for each month. Add appropriate axis labels and a title.

```{r}
# your code here

```

\

Q8: Create a segmented bar chart of the proportion of days with `Ozone` levels above 50 by `Month`. Remove NA values in the `Ozone` column. Add appropriate axis labels and a title.

```{r}
# your code here

```

\

Q9: Create a stacked bar chart showing the count of days available for each month. Fill bar colors by `Ozone` levels above or below 50 for each month. Remove NA values in the `Ozone` column. Add appropriate axis labels and a title.

```{r}
# your code here

```

\

Q10: What is the utility of using a segmented bar chart vs a regular bar chart in this context? Which do you think is more informative and why?

...
